% This is a modified version of Springer's LNCS template suitable for anonymized MICCAI 2025 main conference submissions. 
% Original file: samplepaper.tex, a sample chapter demonstrating the LLNCS macro package for Springer Computer Science proceedings; Version 2.21 of 2022/01/12

\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encodings may result in incorrect characters.
%
\usepackage{graphicx,verbatim}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{tablefootnote}
\usepackage{amssymb}
\usepackage[table]{xcolor}
\usepackage{pifont}
\newcommand{\cmark}{\ding{51}}
\newcommand{\xmark}{\ding{55}}

% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%


%%% my packages
\usepackage{cite}
\usepackage{orcidlink}

\makeatletter
\renewcommand*{\@fnsymbol}[1]{\textbf{\normalsize *}}
\newcommand{\printfnsymbol}[1]{%
  \textsuperscript{\@fnsymbol{#1}}%
}
\makeatother


\begin{document}
%
\title{Exploiting Longitudinal Context in Clinician-Verified Interactive Lesion Tracking}
\titlerunning{Longitudinal Context in Verified Lesion Tracking}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Yannick Kirchhoff\inst{1,2,3}\thanks{Contributed equally. Each co-first author may list themselves as lead on their CV.} \and
Maximilian Rokuss\inst{1,2,3}\printfnsymbol{1} \and
Daniel Philipp Mertens\inst{5} \and
David Füller\inst{5} \and
Benjamin Hamm\inst{1,4} \and
Andreas Schreyer\inst{5} \and
Oliver Ritter\inst{5} \and
Klaus Maier-Hein\inst{1,4,6}
}
%
\authorrunning{Y. Kirchhoff, M. Rokuss et al.}
%
\institute{German Cancer Research Center (DKFZ) Heidelberg, Division of Medical Image Computing, Germany
\and
Faculty of Mathematics and Computer Science, Heidelberg University, Germany
\and
HIDSS4Health -- Helmholtz Information and Data Science School for Health, Karlsruhe/Heidelberg, Germany
\and
Medical Faculty, Heidelberg University, Germany
\and
University Hospital Brandenburg an der Havel, Brandenburg Medical School Theodor Fontane, Germany
\and
Pattern Analysis and Learning Group, Department of Radiation Oncology, Heidelberg University Hospital, Germany\\
\email{\{yannick.kirchhoff,maximilian.rokuss\}@dkfz-heidelberg.de}
}
  
\maketitle              % typeset the header of the contribution
%
% \begin{abstract}
% Tracking tumor lesions across serial CT scans is essential for assessing treatment response in oncology. Current automated methods typically adopt one of two strategies: end-to-end trackers that automatically propagate annotations from baseline to follow-up scans, or decoupled pipelines that first localize the lesion via registration, then segment it independently. The former achieves high automation but lacks clinical safety — if the model tracks an incorrect structure, there is no opportunity for correction. The latter allows user verification of the tracked location but discards the longitudinal context (the lesion's prior appearance), limiting segmentation accuracy in ambiguous cases.
% In this work, we argue for a \textit{Verified Tracking} paradigm, aligned with the autoPET IV challenge setting, where a clinician can but not must correct a propagated prompt before segmentation. We present a unified framework that leverages this verified prompt alongside the visual appearance of the baseline lesion to perform longitudinally-informed segmentation. To address data scarcity, we leverage large-scale synthetic pretraining; this proves essential for exploiting longitudinal context, improving performance by up to 6 Dice points over training from scratch.
% Our approach achieved the top performance in the autoPET IV Challenge (Task 2). Furthermore, to demonstrate generalization beyond the challenge domain, we benchmark on \textit{PanTrack}, a newly curated longitudinal pancreas dataset, which we publicly release to foster further research. Experiments show that incorporating verified prompts with longitudinal context outperforms both single-timepoint baselines and unverified tracking, offering a clinically safe middle ground between automation and control. Code, model and dataset willl be made publicly available at \url{https://github.com/anonymous}.

% \keywords{Longitudinal Tracking \and Interactive Segmentation \and Lesions}
% \end{abstract}

\begin{abstract}
Tracking tumor lesions across serial CT scans is essential for oncological response assessment. Existing automated methods face a fundamental trade-off: end-to-end trackers achieve high automation but offer no opportunity to correct silent tracking failures, while decoupled registration–segmentation pipelines permit user verification yet discard the lesion's prior appearance, limiting accuracy in ambiguous cases. In this work, we propose a \textit{Verified Tracking} paradigm: a clinician verifies a registration-proposed prompt, which the model leverages alongside the baseline lesion appearance to resolve segmentation ambiguities. We present a unified framework combining early spatial prompt fusion with latent temporal difference weighting for longitudinally-informed segmentation. To address data scarcity, we leverage large-scale synthetic pretraining, proving essential for exploiting longitudinal context, improving performance by up to 4.5 Dice points over training from scratch. Our approach secured first place in the MICCAI autoPET IV challenge. We further curate and release \textit{PanTrack,} a new longitudinal pancreatic cancer benchmark, to assess out-of-distribution generalization. Experiments show that our model outperforms prior work in both fully automatic and the proposed verified tracking setting offering a clinically safe middle ground between automation and control. Code, model and dataset will be released at \url{https://github.com/MIC-DKFZ/LongiSeg}.

\keywords{Longitudinal Tracking \and Interactive Segmentation \and Tumor Lesions \and Synthetic Pretraining }
\end{abstract}

\input{chapters/introduction}
% \input{chapters/related_work}
\input{chapters/method}
\input{chapters/datasets}
\input{chapters/experiments_and_results}
\input{chapters/conclusion}

\begin{credits}
\subsubsection{\ackname} This work was partly funded by the Helmholtz Information and Datascience School (HIDSS) and Helmholtz Imaging (HI), platforms of the Helmholtz Incubator on Information and Data Science. Supported by the Helmholtz Foundation Model Initiative (HFMI) through the pilot project THRP (The Human Radiome Project). Funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) – 402688427. This project was funded within the DKTK Heidelberg Seed Funding 25 program. M.R. is funded through a Google PhD Fellowship.

\subsubsection{\discintname} The authors have no competing interests to declare. 
\end{credits}

%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{main}
%
\end{document}
